---
title: "Data Science Challenge"
author: "Luis Carrasco Tornero"
date: "June 3, 2018"
output: rmarkdown::github_document

---

```{r include = FALSE}
library(rmarkdown)
library(rjson)
library(dplyr)
library(ggplot2)
```


## Part 1: Experiment design

__AirBnB Professional photography: 1. Provide full details about how you will run experiments to assess the impact of this service on both hosts and guests. How will you ensure that the experiments are valid and not biased?__

Two different approaches could be used to test whether professional photography services are impacting the success of the properties.

1) Experiment based on the comparison between properties using the services against those not using the service.

2) Temporal comparison, analyzing the success increase (or not) of properties that adopted the professional photography but did not used it in the past.

For __experiment 1)__, data of both hosts and guests will be compiled. Examples of the data compiled would be:

- Host data: Number of visits, number of bookings, price/night/guest, city, location, property characteristics (number of bedrooms, bed size, etc.).

- Guest data: Number of visits, type of booked host, time of each visit, rating of host description 'accuracy'.

The ratio visits/booking would be compared between hosts with and without photography service. Time of each visit for guests of hosts with and without the service would be compared. Also the ratio of bookings for hosts with and without service will be compared for similar type hosts (see avoiding biases section). Comparison of the 'accuracy' rating would be done. Stats: The analysis could be performed using tests for equality of means. 

__Experiment 2)__ would analyze success changes of properties changing from not using service to using one (and the opposite direction, if there are cases). This experiment would analyze hosts (using the data of experiment 1). The price of the property, the number of bookings per unit of time, as well as the ratio visits/bookings will be compared from before and after using the service. Increases in the price after adopting the service would be especially useful. 
Stats: Step functions (in comparison with continuous functions) and time-series analysis based (i.e. change-point detection) methods could be used.


__Avoiding biases:__ Comparison between photography service users and not users is not trivial because the booking success is affected by many variables, such as the price or quality of the property. In order to avoid those biases, separate analysis controlling for those variables are needed. This are a few examples of how to avoid this biases:

- Use data only from 2015. The launch of identity verification, double blind reviews, increase of PR for AirBnB and the high quality of smart phones cameras could by decreasing the relative importance of the professional photography. Using data from before 2015 could give an overestimated value of the service.

- Compare data for same price ranges: More expensive hosts could be more dependent on high quality photography, as people staying at pricey properties might be looking for beauty. Guests looking for cheaper properties might not be looking for pretty-looking properties but for convenient ones.

- Compare data for similar locations: Very busy locations could be less influenced by the quality of the photography, as visitors look for cheaper deals no matter of the looks of the property. On the other hand, countryside locations could be relying on the looks.

- Comparing by price and location might not be enough to avoid quality bias, as some properties might have better facilities. Number of rooms, bathrooms, room size or facilities could be compared. Some index could be built to compile all this characteristics into a 'quality' measure. The analysis should compare properties with similar quality.

Considering all these biases would be fundamental, as the results could show that the best strategy might be to offer the service just to expensive properties (the service is more important and the benefit for AirBnB is bigger), or to properties located in the countryside. 


## Part 2: Result analysis

_Pick-up time optimization challenge_

In order to the answer the questions proposed in the challenge: "should the company move towards _road_ distance?", the main purpose of the data analysis is to clarify whether or not the duration of the rides to pick up the customers are shorter for the _road_ vehicles than for the _linear_ vehicles.

To do that, the focus will be on the _going_to_pickup_ data, and the mean values for the duration of the trips will be analysed for different parameters.

First, the data is extracted:

```{r, message=FALSE, warning=FALSE}
json_file <- "intervals_challenge.json"
data <- fromJSON(sprintf("[%s]",paste(readLines(json_file),collapse = ",")))
df_raw <- do.call(rbind.data.frame,data)
print(paste("Number of data elements = ",nrow(df_raw)*ncol(df_raw),sep=""))

```

The data contains more than a million elements. The size of the data could represent slow computing times so some data cleansing is needed. The purpose of the challenge is to evaluate whether the new _road_ assignment will reduce the pick up duration, so the other intervals data are not needed. Other variables, such as distance or vehicles ids are not needed for our analysis either. 

```{r}
# Subset to "going_to_pickup"
df_sub1 <- subset(df_raw,type=="going_to_pickup")
# Exclude non-used variables
vars <- names(df_sub1) %in% c("distance", "vehicle_id", "type")
df_sub <- df_sub1[!vars]
# Remove non-desired data
rm(data,df_raw,df_sub1)

print(paste("Number of data elements = ",nrow(df_sub)*ncol(df_sub),sep=""))
```

Now the number of elements have been reduce more than 4 times.

The *trip_id* needs to be reclassified into _road_ and _linear_ factors in order to be analysed:

```{r}
# Assign to "road"
df_sub$trip_id <-  gsub("^0\\w+|^1\\w+|^2\\w+|^3\\w+|^4\\w+|^5\\w+|^6\\w+|^7\\w+|^8\\w+", 
                        "road", df_sub$trip_id)
# Assign to "linear"
df_sub$trip_id <- gsub("^9\\w+|^a\\w+|^b\\w+|^c\\w+|^d\\w+|^e\\w+|^f\\w+", 
                        "linear", df_sub$trip_id)
```


Now we want to see how the data looks like without controlling for city, time of the day, or ride duration range. To do that, boxplots for each assignment (*road* and _linear_) showing the median and quartiles are used:


```{r}
p1 <- ggplot(aes(y = as.numeric(duration), x = as.factor(trip_id)), 
             data = df_sub) + geom_boxplot()
p1 + xlab("Assigment") + ylab("Duration (s)") 
```

Median values look virtually identical. But to test whether _linear_ and _road_ assignment is providing different duration, a statistical test is needed. To choose the proper test to be used for our data, we need to first test whether the data is normally distributed. A Shaphiro-Wilk normality test for _duration_ is used for each assignment:

```{r}
# Shapiro-Wilk normality test for road duration
with(df_sub[sample(1:nrow(df_sub),5000),], 
     shapiro.test(as.numeric(duration[trip_id == "linear"])))
# Shapiro-Wilk normality test for linear duration
with(df_sub[sample(1:nrow(df_sub),5000),], 
     shapiro.test(as.numeric(duration[trip_id == "road"])))
```

The W values and p-values indicate that the data is not normally distributed. Because of this, t-tests cannot be used. We then have to run a Wilcoxon test for non-normal distributed data in order to test the equality of the mean values of _duration_.

```{r}
wilcox.test(as.numeric(df_sub$duration[df_sub$trip_id=="road"]), 
            as.numeric(df_sub$duration[df_sub$trip_id=="linear"]), 
            alternative = "two.sided")
```

The Wilcoxon test indicates that we can not assume that the means are different. In other words, the means of _duration_ are not different, and there is no effect of the assignment type on the duration of the rides.

We can compare the mean duration values for _linear_ and _road_:

```{r}
mean(as.numeric(df_sub$duration[df_sub$trip_id=="linear"]))
mean(as.numeric(df_sub$duration[df_sub$trip_id=="road"]))
```

The _road_ assignment was, in average, only 2 seconds faster. Although this is not statistically significant and this difference could be due to randomness.


Next, the same analysis was performed for different intervals of duration time. One could argue that the _normal_ assignment could work better for larger distances (and therefore larger duration), so that the mean value considering all durations is not taking into account some particular duration times. Larger distances mean more uncertainty and therefore _normal_ could work better. 

To test this, an analysis was done for different duration intervals with same results. Preliminary analysis showed smaller duration for the _road_ assignment when the duration was large (distant places), and no differences for small durations. For example, this is the difference for durations above 3000 seconds:

```{r}
p2 <- ggplot(aes(y = as.numeric(duration), x = as.factor(trip_id)), 
             data = df_sub[as.numeric(df_sub$duration)>3000,]) + geom_boxplot()
p2 + xlab("Assigment") + ylab("Duration (s)") 
```

Slightly shorter median duration is observed for the _road_ assignment, but not significant.


Next, one could wonder whether these results are city-dependent. Bigger and complex cities could be benefited by _road_ as traffic assessments could help decide the best routes, while smaller cities could not show the same benefits.

To test this, we repeated the analysis but controlling by city:

```{r}
p3 <- ggplot(aes(y = as.numeric(duration), x = as.factor(city_id), 
                 fill = as.factor(trip_id)), data = df_sub) + geom_boxplot() 
p3 + xlab("City") + ylab("Duration (s)") + theme(legend.title=element_blank()) 
```

Volantis seems to be benefited by the _road_ assignment, but a formal test for the equality of the means must be performed to be able to infer conclusions. The same methodology was used as in previous sections. First, the normality of the distribution was checked. As the distribution of the data was not normal, a Wilcoxon test was used.

```{r, echo=T, results='hide'}
df_bravos <- df_sub[df_sub$city_id=="bravos",]
df_pentos <- df_sub[df_sub$city_id=="pentos",]
df_volantis <- df_sub[df_sub$city_id=="volantis",]
# Normality test
with(df_bravos[sample(1:nrow(df_bravos),5000),], 
     shapiro.test(as.numeric(duration[trip_id == "linear"]))) #p-value < 2.2e-16
with(df_pentos[sample(1:nrow(df_pentos),5000),], 
     shapiro.test(as.numeric(duration[trip_id == "linear"]))) #p-value < 2.2e-16
with(df_volantis[sample(1:nrow(df_volantis),5000),], 
     shapiro.test(as.numeric(duration[trip_id == "linear"]))) #p-value < 2.2e-16

# Equality of means test
wilcox.test(as.numeric(df_bravos$duration[df_bravos$trip_id=="road"]),
            as.numeric(df_bravos$duration[df_bravos$trip_id=="linear"]), 
            alternative = "two.sided")                        #p-value = 0.5828
wilcox.test(as.numeric(df_pentos$duration[df_pentos$trip_id=="road"]),
            as.numeric(df_pentos$duration[df_pentos$trip_id=="linear"]), 
            alternative = "two.sided")                        #p-value = 0.9311
wilcox.test(as.numeric(df_volantis$duration[df_volantis$trip_id=="road"]),
            as.numeric(df_volantis$duration[df_volantis$trip_id=="linear"]), 
            alternative = "two.sided")                        #p-value = 0.3462
```

The W and p-values show that it cannot be concluded that the means are different. The duration of the travel of each assignment is equal for each city as well.


Finally, one could think there could exist differences between _road_ and _linear_ for different times of the day. For example, one could think than during peak hours (getting/going back from work) the _normal_ assignment could mean faster trips due to the capability of predicting traffic jams.
To check this, a similar analysis to previous parts was done. The duration data was grouped into 12 intervals, by dividing the started_at values into 12 groups.

We create the one boxplot for each time interval (the whole time series divided by 12):

```{r}
# Start_time is broken into 12 intervals
df_sub$tinterv <- as.numeric(cut(df_sub$started_at, breaks = 12))
# Boxplots are plotted for each time interval
p4 <- ggplot(aes(y = as.numeric(duration), x = as.factor(tinterv), fill = 
                   as.factor(trip_id)), data = df_sub) + geom_boxplot() 
p4 + xlab("Day interval") + ylab("Duration (s)") + theme(legend.title=element_blank()) 
```

We can see that the differences between _road_ and _linear_ are minimal for most of the time intervals. The time interval with a higher difference was the interval 7, but in this case the _linear_ durations were shorter than those of the _road_ assignment.



**1. Should the company move towards road distance? What's the max price it would make sense to pay per query? (make all the assumptions you need, and make them explicit)**

The analysis showed that the trips of the vehicles that were picking up customers did not differ in duration
from the group that used the _linear_ assignment than the group that used _road_ assignment. We can conclude from this that the company should not move towards the road distance. The _road_ service comes with certain cost and increases the complexity of the system, potentially decreasing its reliability. After observing that the use of this service does not translate into faster trips to picking up costumers, it is safe to say that it should not be used.

A maximum price for each query of the _road_ service was not calculated as any value bigger than zero would not be worth it, because the analysis showed that the difference in duration from the two systems can be due just to randomness and not an improvement of the speed due to the _road_ service.

A different question would be if the _road_ system improves the gas consumption due to avoiding less energy-efficient routes. For example, _road_ takes the same time to arrive to destination, but through a route free of traffic lights were the gas consumption is reduced. In this case, _road_ system would have benefits. But because the question was related to the benefits for the time-reducement to pick up to the costumer, the answer is that the company should not move towards road distance.

**2. How would you improve the experimental design? Would you collect any additional data?**

There are a few aspects that could be improved from the experimental design. First, the experiment has been done in three cities with very similar duration times. This means that the size of the city, or the density of drivers is similar between them (although the time duration for Bravos was significantly higher than the other two). To obtain less biased results, a wide range of cities presenting a range of durations (which would mean a range of city size or drivers density) would be needed. This could show, for example, that for cities with lower density of drivers, the _road_ system might have some advantage.

An interesting additional data to collect during the experiment would have been the driver's coordinates (initial and final). Making the problem spatially explicit could have provided more insides of the advantages/disadvantages of each system. For example, adding coordinates would have allowed to make a study based in regions. In more populated areas, or with more costumers, or with higher traffic densities, differences between _road_ and _linear_ systems could have arised.